---
title: "ST558_Project1_LanLin"
author: "Lan Lin"
date: '2022-06-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'F:\\Graduate\\NCSU_courses\\ST558\\projects\\project_1')
```

```{r, include=FALSE}
library(easypackages)
libraries("httr", "tidyverse", "jsonlite", "DT", "stringr", "RColorBrewer")
```


# Part 1: Required packages 
### To use the functions for interacting with the Covid API, I used the following packages:

`easypackage`: Easily load and install multiple packages from different sources 
`httr`: Useful tools for working with HTTP organised by HTTP verbs
`tidyverse`: tons of useful features for data manipulation and visualization
`jsonlite`: API interaction
`DT`: Provides an R interface to the JavaScript library DataTables
`stringr`: Provides a cohesive set of functions designed to make working with strings as easy as possible. 
`RColorBrewer`: Uses the work from http://colorbrewer2.org/ to help us choose sensible color schemes for figures in R

# Part 2: API Interaction Functions

### 2.1 Write a function to read in the data set from API.

All queries are prefixed with https://api.covid19api.com/ and are GET requests unless otherwise noted.
```{r}
# Write a function to read in the data set from API
get_Data <- function(url){
                myData<- GET(paste0("https://api.covid19api.com", url))
                parsed <- fromJSON(rawToChar(myData$content))
                return(parsed)
}
```

### 2.2 Write a function to rename the countries to match the names in map data set.
```{r}
countries_Names <- function(countries){
  ifelse(countries %in% c("Antigua and Barbuda", "Brunei Darussalam", "Iran, Islamic Republic of", "Micronesia, Federated States of", "Taiwan, Republic of China", "Tanzania, United Republic of", "Trinidad and Tobago"), new_countries <- str_replace(word(countries, 1), ",", ""), 
         ifelse(countries == "Congo (Brazzaville)",  new_countries <- "Republic of Congo", 
                ifelse(countries == "Congo (Kinshasa)",  new_countries <- "Democratic Republic of the Congo",
                       ifelse(countries == "Korea (North)", new_countries <- "North Korea",
                              ifelse(countries == "Lao PDR", new_countries <- "Laos",
                                     ifelse(countries == "Macedonia, Republic of", new_countries <- "North Macedonia", 
                                            ifelse(countries == "Republic of Kosovo", new_countries <- "Kosovo", 
                                                  ifelse(countries == "Saint Vincent and Grenadines", new_countries <- "Saint Vincent",
                                                         ifelse(countries == "Syrian Arab Republic (Syria)", new_countries <- "Syria",
                                                                ifelse(countries == "United Kingdom", new_countries <- "UK",
                                                                       ifelse(countries == "United States of America", new_countries <- "USA",
                                                                              ifelse(countries == "Viet Nam", new_countries <-"Vietnam", 
                                                                                     ifelse(countries == "Saint Kitts and Nevis", new_countries <- "Saint Kitts", 
                                                                                            ifelse(countries == "Korea (South)", new_countries <- "South Korea",
                                                                                                   ifelse(countries == "Russian Federation", new_countries <- "Russia", 
                                                                                                          new_countries <- countries)))))))))))))))
}
```

### 2.3 Write a function to rename the countries in the population data set to match the names in the API data sets.
```{r}
countries_Names_pop <- function(countries){
  ifelse(countries == "Antigua And Barbuda", new_Countries <- "Antigua and Barbuda",
         ifelse(countries == "Bosnia And Herzegovina", new_Countries <- "Bosnia and Herzegovina",
                ifelse(countries == "Czechia", new_Countries <- "Czech Republic",
                       ifelse(countries == "Democratic Republic Of The Congo", new_Countries <- "Congo (Kinshasa)",
                              ifelse(countries == "Congo", new_Countries <- "Congo (Brazzaville)",
                                     ifelse(countries == "Guinea Bissau", new_Countries <- "Guinea-Bissau",
                                            ifelse(countries == "Holy See", new_Countries <- "Holy See (Vatican City State)",
                                                   ifelse(countries == "Iran", new_Countries <- "Iran, Islamic Republic of", 
                                                          ifelse(countries == "South Korea", new_Countries <- "Korea (South)",
                                                                 ifelse(countries == "Laos", new_Countries <- "Lao PDR", 
                                                                        ifelse(countries == "North Macedonia", new_Countries <- "Macedonia, Republic of",
                                                                               ifelse(countries == "Micronesia (Federated States Of)", new_Countries <- "Micronesia, Federated States of", 
                     ifelse(countries == "Palestine", new_Countries <- "Palestinian Territory", 
                            ifelse(countries == "Kosovo", new_Countries <- "Republic of Kosovo",
                                   ifelse(countries == "Russia", new_Countries <- "Russian Federation",
                                          ifelse(countries == "Saint Kitts And Nevis", new_Countries <- "Saint Kitts and Nevis",
                                                 ifelse(countries == "Sao Tome And Principe", new_Countries <- "Sao Tome and Principe",
                                                        ifelse(countries == "Syria", new_Countries <- "Syrian Arab Republic (Syria)",
                                                               ifelse(countries == "Taiwan", new_Countries <- "Taiwan, Republic of China",
                                                                      ifelse(countries == "United Republic Of Tanzania", new_Countries <- "Tanzania, United Republic of",
                      ifelse(countries == "Timor Leste", new_Countries <- "Timor-Leste",
                             ifelse(countries == "Trinidad And Tobago", new_Countries <- "Trinidad and Tobago",
                                    ifelse(countries == "United States Of America", new_Countries <- "United States of America",
                                           ifelse(countries == "Venezuela", new_Countries <- "Venezuela (Bolivarian Republic)",
                                                  ifelse(countries == "Vietnam", new_Countries <- "Viet Nam", new_countries <- countries)))))))))))))))))))))))))
}
```

### 2.4 Write a function to create a week list based on start date and end date.
```{r}
get_Date <- function(start, end){
  diff <- as.numeric(difftime(as.Date(end),as.Date(start)))
  start_date <- seq(as.Date(start), by = "week", length.out = diff %/% 7 +1)
  end_date <- c(seq(as.Date(start)+6, by = "week", length.out = diff %/% 7), end)
  return(list(start_date, end_date))
}
```

### 2.5 Write a function to get the confirmed cases based on start date and end date of any country.
```{r}
get_confirmed_data <- function(start, end, country){
  Week_list<-get_Date(start, end)
  web_url_1<-paste0("/country/","united-states","/status/confirmed?")
  country_data_total = data.frame()
  for(i in 1:length(Week_list[[1]])){
    web_url_2<-paste0(web_url_1,"from=",Week_list[[1]][[i]],"T00:00:00Z","&to=",Week_list[[2]][[i]],"T00:00:00Z")
    country_data <-get_Data(web_url_2)
    country_data_total<-rbind(country_data_total,country_data)
  }
  return (country_data_total)
}
```

# Part 3: Data manipulation 

### 3.1 Read in the summary data set on current day. 
This data set includes the numbers of new deaths, total deaths, new recovered, total recovered of current day worldwide. 

```{r}
# Read in the data set
covid_summary <- get_Data("/summary")$Countries %>% as_tibble()

# Using the countries_Names function to create a new variable called new_Countries to match the country names in the map data set
covid_summary$new_Countries <- countries_Names(covid_summary$Country)
```

### 3.2 Read in the map data set.
```{r}
# Read in the world map
mapdata <- map_data("world")

# Combine the map and covid_summary data sets 
mapData <- left_join(mapdata, covid_summary, by = c("region" = "new_Countries"))

# Remove the rows which do not contain total confirmed cases numbers 
mapdata1 <- mapData %>% filter(!is.na(mapData$TotalConfirmed))
```


### 3.3 Read in the population data set. 
This is data set is downloaded from "https://www.ecdc.europa.eu/en/geographical-distribution-2019-ncov-cases".
```{r, include=FALSE}
# Read in the data set
pop <- distinct(read_csv("data.csv")[1:4])

# Using the countries_Names_pop function to create a new variable Country to match the country names in the covid_summary data set. 
pop$Country <- countries_Names_pop(pop$country)
pop <- pop[3:5]

# Combine the covid_summary and population data sets 
covid_summary_pop <- left_join(covid_summary, pop, by = "Country")
```

### 3.4 Create a tibble to plot histogram 
```{r}
# Create a new variable which represents the proportion of population who contrasted the corona virus 
covid_summary_hist <- covid_summary_pop %>% 
                      mutate(ConfirmedProportion = TotalConfirmed/population) %>%
                      select(Country, ConfirmedProportion) %>%
                      drop_na(ConfirmedProportion)
```

### 3.5 Read in the US confirmed cases information in the period of Jan 01, 20222 to May 31, 2022
```{r}
# Read in the data set
US_Data <- get_confirmed_data("2022-01-01","2022-05-31","united-states")
# Check the data quality
summary(US_Data$Cases)
```

The minimum total confirmed cases is -3073, which means that some data entries were wrong. Since we are not able to fix the data, I just removed the rows which had cases less than 0. 

```{r}
# Remove the rows which had cases less than 0, and create a new variable call NewCases
US_NewCase <- US_Data %>% 
              filter(Cases >= 0) %>%
              group_by(Province, City) %>%
              select(Cases, Date) %>%
              mutate(NewCases = Cases - lag(Cases, default = first(Cases))) %>%
              filter(Province != "" & NewCases >= 0)
```
```{r}
# Create a tibble to summary the total new cases by state
US_NewCase_StateSum <- US_NewCase %>%
                      group_by(Province) %>% 
                      summarise(TotalNewCases = sum(NewCases))

state_names <- pull(US_NewCase_StateSum %>% arrange(desc(TotalNewCases)) %>% top_n(10) %>% select(Province))

# Create a tibble to the new case information in the top 10 states 
US_NewCase_byState <- US_NewCase %>% group_by(Province, Date) %>% summarise(Sum = sum(NewCases)) %>% filter(Province %in% state_names)
```



# Part 4: EDA 

### 4.1 Plot a spatial data to show the coronavirus cases in the world 
```{r}
ggplot(mapdata1, aes(x= long, y =lat, group = group)) +
  geom_polygon(aes(fill = TotalConfirmed), color = "darkgreen") + 
  scale_fill_gradient(name = "Cases", low='#EEEEEE', high='darkgreen', na.value = "grey") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        rect = element_blank()) +
  labs(title = "Coronavirus Worldwide")

# labels <- covid_summary %>% arrange(desc(TotalConfirmed)) %>% select(new_Countries) %>% slice(1:10)
```
This map shows us that the United States of America has the largest total confirmed cases, followed by India. In order to further demonstrate the corona virus cases by country, two tables are created. 

### 4.2 Summary corona virus cases by country
```{r}
case_tab <- covid_summary_pop %>% 
              select(c("Country", "continent", "NewConfirmed", "TotalConfirmed", "NewDeaths", "TotalDeaths", "population"))  %>%
                arrange(desc(TotalConfirmed)) %>%
                  rename(Continent = continent, Population = population)
datatable(case_tab)
```

### 4.3 Summary corona virus cases by continent 
```{r} 
case_summary_tab <- covid_summary_pop %>% 
                      drop_na(continent) %>%  
                        group_by(continent) %>% 
                          summarise(ContinentTotalConfirmed = sum(TotalConfirmed), ContinentTotalDeaths =sum(TotalDeaths)) %>%
                            arrange(desc(ContinentTotalConfirmed))

row_names <- case_summary_tab$continent

case_summary_tab <- rbind(case_summary_tab[2:3],colSums(case_summary_tab[2:3]))
case_summary_tab <- as.data.frame(case_summary_tab)

rownames(case_summary_tab) <- c(row_names, "Total")

datatable(case_summary_tab)
```

This table shows us that Europe has the largest total confirmed cases, while America has the largest total deaths. The total confirmed cases and total deaths are 540740012	and 6318769, respectively. 

### 4.4 Create numerical summaries for total confirmed cases and deaths in different contients 
```{r}
#  Create numerical summaries for total confirmed cases and deaths in different consents
tapply(X= c(covid_summary_pop$TotalConfirmed), INDEX = covid_summary_pop$continent, FUN = summary)
```
This table summarize the minimum, first quantile, median, mean, third quantile and maximum total confirmed cases in each continent. And the America has the most confirmed cases among all the consents. 

```{r}
#  Create numerical summaries for total deaths in different consents
tapply(X= c(covid_summary_pop$TotalDeaths), INDEX = covid_summary_pop$continent, FUN = summary)
```
This table summarize the minimum, first quantile, median, mean, third quantile and maximum total deaths in each continent. And the America has the most deaths among all the consents. 

### 4.5 Create a histogram to show the distribution of the proportion of population who contrasted the corona virus by country 
```{r}
ggplot(covid_summary_hist, aes(x= ConfirmedProportion)) +
  geom_histogram(binwidth = 0.02, fill="#69b3a2") +
  labs(title = "Histogram Plot of the Proportion of Confirmed Cases")
```
This histogram shows us that most of countries had less than 20% of their population contracted the corona virus. This proportions were up to roughly 60% in some countries. However, this numbers are highly depending on the proportion of people who had covid test. 

### 4.6 Create a scatter plot to show the relationship between total confirmed cases and total deaths by country
```{r, warning= FALSE, fig.width= 15, fig.width= 15}
confirm_vs_death <-ggplot(covid_summary_pop, aes(x = TotalConfirmed, y = TotalDeaths))

confirm_vs_death + 
  geom_point() +
  scale_x_continuous(trans='log2') + scale_y_continuous(trans='log2') +
  geom_text(aes(label = Country, color = factor(continent)), size = 3.5, check_overlap = TRUE, vjust = "inward", hjust = "inward") +
  labs(title = "Scatter Plot of the relationship between total confirmed cases and total deaths by country")
```
In this scatter plot, I did log2 transformation of both TotalConfrimed and TotalDeath. This plot shows us that log2(TotalConfrimed) and log2(TotalDeath) had a roughly linear relationship among countries. The number of total deaths increases as the number of total confirmed cases increases. North Korea, Antarctica and Micronesia are outlines in this plot. This may be due to the low mortality rate in thes countries or lack of data. 

Since the United States had both the highest total confirmed cases and total deaths, I would like to look into some details of US in the period Jan 1, 2022 to May 31, 2022. 

### 4.7 Create a table to summary the total new cases in the period Jan 1, 2022 to May 31, 2022 by state. 
```{r}
datatable(US_NewCase_StateSum %>% arrange(desc(TotalNewCases)))
```

### 4.8 Create a box plot of Daily New Cases in the top 10 States in the period of Jan 1, 2022 to May 31, 2022 by state. 
```{r}
ggplot(US_NewCase_byState, aes(x = Province, y = Sum)) +
  geom_boxplot() + 
  coord_cartesian(ylim=c(0, 75000)) +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(x= "Sates", y= "Daily New Cases", title = "Box Plot of Daily New Cases in the top 10 States")
```

This box plot shows us that California had the largest range and highest daily new cases in the period of Jan 1, 2022 to May 31, 2022. And all the top 10 states had a lot of outliners. 


```{r}
a <- US_NewCase %>% filter(Province == "California" & NewCases >= 5000)
table(a$City)
```













